---
title: "Into to sf package"
author: "Anthony Twesigye"
date: "10/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



## Intro


### Load data & Libraries

```{r load_libs_data}

library(sf)
library(tidyverse)

# # get default directory
# current_wd <- getwd()
# # change to the working directory
# setwd("2021-05-10")
admin <-  sf::st_read(dsn = "inputs", layer = "uganda_regions")
admin
# # reset to default directory
# setwd(current_wd)

```

### visualize sf

```{r ggplot_sf}
ggplot()+
  geom_sf(data = admin)

```

### Create some sample point data

```{r }
pt_sample <- st_sample(x = admin, size = 600)
class(pt_sample)
class(admin)

ggplot()+
  geom_sf(data = admin)+
  geom_sf(data = pt_sample)

```

### spatial joints
- pt_did not have data frame. Used st_as_sf to give data frame which allows spatial join
```{r }
pt_sample <- pt_sample %>% 
  st_as_sf() %>% 
  st_join(admin) %>% 
  select(Region)

```

### spatial verification
- rename pt sample to target_sample, add pt id
- create hh interview by jittering 150m, removing pt id, and randomly shuffling rows

```{r }
target_sample <- pt_sample %>% 
  mutate(pt_id = row_number())

hh_interviews <- target_sample %>% 
  st_jitter(amount = 150) %>% 
  select(-pt_id) %>% 
  sample_n(nrow(.))

```

### Find closest point
- create index to closest feature
```{r }
target_sample_index <- hh_interviews %>% 
  st_nearest_feature(target_sample)

closest_pt_distances <- hh_interviews %>% 
  st_distance(target_sample[target_sample_index,], by_element = TRUE)
closest_pt_distances[1:20]

```

- first create enumerator_id to mimic assessment data
```{r}
hh_interviews %>% 
  mutate(enumerator_id = rep(1:10,60))

```

- build a function which combines these steps

```{r, distance_function}

mutate_nearest_feature <- function(x, y) {
  y_index <- x %>% 
    st_nearest_feature(y)
  
  closest_distance <- x %>% 
    st_distance(y[y_index,], by_element = TRUE)
  
  x %>% 
    mutate(
      y_index = y_index,
      closest_distance = closest_distance
    )
  
}

```

- test out function

```{r run_function}
mutate_nearest_feature(x=hh_interviews, y = target_sample)

hh_w_target <- hh_interviews %>% 
  mutate_nearest_feature(target_sample)

# hh_w_target <- hh_interviews %>% 
#   butteR::mutate_nearest_feature(target_sample)

```


- filter to entries greater than or equal to 150m
```{r }
hh_w_target %>%
  filter(as.numeric(closest_distance) >= 150)

```


### find distance to target point by point ID

```{r}

hh_interviews <- target_sample %>% 
  st_jitter(amount = 150) %>% 
  mutate(enumeratoer_id = rep(1:10, 60)) %>% 
  sample_n(nrow(.))

pt_id_index <- hh_interviews$pt_id
distance_to_target <-  st_distance(x=hh_interviews, y = target_sample[pt_id_index,], by_element = TRUE)

distance_to_target[1:10]

```

### Task 1: Turn above code into function to measure distance to target ID using above code
- Need some clarity on the filtering with indices indices.

```{r dist_to_target}
distance_to_target_id  <- function(input_x, input_y) {
  
  pt_id_index <- input_x$pt_id
  distance_to_target <-  st_distance(x=input_x, y = input_y[input_y$pt_id==pt_id_index,], by_element = TRUE)
  
  input_x %>% 
    mutate(
      pt_id_index = pt_id_index,
      distance_to_target = distance_to_target
    )
  
}

```

- Run Task 1
```{r run_task_1}
# random hh_inteviews
hh_interviews <- target_sample %>% 
  st_jitter(amount = 150) %>% 
  mutate(enumeratoer_id = rep(1:10, 60)) %>% 
  sample_n(nrow(.)) %>% 
  arrange(pt_id)

# run the function
task_1_output <- distance_to_target_id(hh_interviews, target_sample)

# check output
task_1_output %>% 
  filter(as.numeric(distance_to_target) >= 150)

```

### Task 2:
