---
title: "Purr"
author: "Anthony Twesigye"
date: "30/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

## https://www.rebeccabarter.com/blog/2019-08-19_purrr/

## Some map functions

### map()

- map(.x, .f) is the main mapping function and returns a list
- map_df(.x, .f) returns a data frame
- map_dbl(.x, .f) returns a numeric (double) vector
- map_chr(.x, .f) returns a character vector
- map_lgl(.x, .f) returns a logical vector

### map2()

### pmap()

```{r}
my_first_list <- list(my_number = 5,
                      my_vector = c("a", "b", "c"),
                      my_dataframe = data.frame(a = 1:3, b = c("q", "b", "z"), c = c("bananas", "are", "so very great")))
my_first_list

```

## Simplest usage: repeated looping with map

### map(.x, .f)

```{r}
addTen <- function(.x) {
  return(.x + 10)
}

# vector
added_values <- map(.x = c(1, 4, 7), .f = addTen)
added_values[[1]]

added_values_2 <- map(c(1, 4, 7), addTen)

# list
map(list(1, 4, 7), addTen)

# data frame
map(data.frame(a = 1, b = 4, c = 7), addTen)

# in case of need of different output
map_dbl(c(1, 4, 7), addTen)
map_chr(list(1, 4, 7), addTen)
map_df(data.frame(a = 1, b = 4, c = 7), addTen)

map_df(c(1, 4, 7), function(.x) {
  return(data.frame(old_number = .x, 
                    new_number = addTen(.x)))
})
# Note that in this case, I defined an “anonymous” function as our output for each iteration.

# using ay argument name
map_df(c(1, 4, 7), function(input) {
  return(data.frame(old_number = input, 
                    new_number = addTen(input)))
})

# modify(), which is just like the map functions, but always returns an object the same type as the input object
modify(c(1, 4, 7), addTen)
modify(list(1, 4, 7), addTen)
modify(data.frame(1, 4, 7), addTen)
# modify_if(), that only applies the function to elements that satisfy a specific criteria (specified by a “predicate function”, the second argument called .p)
modify_if(.x = list(1, 4, 7), 
          .p = function(x) x > 5,
          .f = addTen)
```

## The tilde-dot shorthand for functions

To make the code more concise you can use the tilde-dot shorthand for anonymous functions (the functions that you create as arguments of other functions).

The notation works by replacing

``` r
function(x) {
  x + 10
}
```

with

``` r
~{.x + 10}

```

`~` indicates that you have started an anonymous function, and the argument of the anonymous function can be referred to using `.x` (or simply `.`)

```{r}
map_dbl(c(1, 4, 7), ~{.x + 10})

```

## gapminder dataset

```{r}
# to download the data directly:
gapminder_orig <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")
# define a copy of the original dataset that we will clean and play with 
gapminder <- gapminder_orig

dim(gapminder)
head(gapminder)

# identify the type of each column
gapminder %>% map_chr(class)

# number of distinct values in each column
gapminder %>% map_dbl(n_distinct)

# return a few different summaries of each column in a data frame, you can use map_df()

gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))))

# you can tell map_df() to include column names using the .id argument of map_df()

gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))),
                     .id = "variable")

```

If you’re having trouble thinking through these map actions, I recommend that you first figure out what the code would be to do what you want for a single element, and then paste it into the map_df() function

For instance, since the first element of the gapminder data frame is the first column, let’s define `.x` in our environment to be this first column.

```{r}
# take the first element of the gapminder data
.x <- gapminder %>% pluck(1)
# look at the first 6 rows
head(.x)

# Then, you can create a data frame for this column that contains the number of distinct entries, and the class of the column
data.frame(n_distinct = n_distinct(.x),
           class = class(.x))
# Since this has done what was expected want for the first column, you can paste this code into the map function using the tilde-dot shorthand.
gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))),
                     .id = "variable")

```

## Maps with multiple input objects

The map function that maps over two objects instead of 1 is called `map2()`. The first two arguments are the two objects you want to iterate over, and the third is the function (with two arguments, one for each object).

``` r
map2(.x = object1, # the first object to iterate over
     .y = object2, # the second object to iterate over
     .f = plotFunction(.x, .y))
```

```{r}
continent_year <- gapminder %>% distinct(continent, year)
continent_year

# extract the continent and year pairs as separate vectors
continents <- continent_year %>% pull(continent) %>% as.character
years <- continent_year %>% pull(year)

```

If you want to use tilde-dot short-hand, the anonymous arguments will be `.x` for the first object being iterated over, and `.y` for the second object being iterated over.

Before jumping straight into the map function, it’s a good idea to first figure out what the code will be for just first iteration (the first continent and the first year, which happen to be Asia in 1952).

```{r}

# try to figure out the code for the first example
.x <- continents[1]
.y <- years[1]
# make a scatterplot of GDP vs life expectancy in all Asian countries for 1952
gapminder %>% 
  filter(continent == .x,
         year == .y) %>%
  ggplot() +
  geom_point(aes(x = gdpPercap, y = lifeExp)) +
  ggtitle(glue::glue(.x, " ", .y))

```

This seems to have worked. So you can then copy-and-paste the code into the `map2` function

```{r}
plot_list <- map2(.x = continents, 
                  .y = years, 
                  .f = ~{
                    gapminder %>% 
                      filter(continent == .x,
                             year == .y) %>%
                      ggplot() +
                      geom_point(aes(x = gdpPercap, y = lifeExp)) +
                      ggtitle(glue::glue(.x, " ", .y))
                  })

plot_list[[1]]

plot_list[[22]]

```

`pmap()` allows you to iterate over an arbitrary number of objects (i.e. more than two)


## List columns and Nested data frames

A tibble can be “nested” where the tibble is essentially split into separate data frames based on a grouping variable, and these separate data frames are stored as entries of a list (that is then stored in the `data` column of the data frame).

```{r}
gapminder_nested <- gapminder %>% 
  group_by(continent) %>% 
  nest()
gapminder_nested

gapminder_nested$data[[1]]

```

Using dplyr `pluck()` function, this can be written as

```{r}
gapminder_nested %>% 
  # extract the first entry from the data column
  pluck("data", 1)

# the 5th entry in the data column corresponds to the entire gapminder dataset for Oceania.
gapminder_nested %>% 
  pluck("data", 5)

```

You now have the power to use dplyr manipulations on more complex objects that can be stored in a list.

However, since actions such as mutate() are applied directly to the entire column (which is usually a vector, so is fine), we run into issues when we try to mutate a list. For instance, since columns are usually vectors, normal vectorized functions work just fine on them

```{r}
tibble(vec_col = 1:10) %>%
  mutate(vec_sum = sum(vec_col))
```

but when the column is a list, vectorized functions don’t know what to do with them, and we get an error that says `Error in sum(x) : invalid 'type' (list) of argument`. Try

```r
tibble(list_col = list(c(1, 5, 7), 
                       5, 
                       c(10, 10, 11))) %>%
  mutate(list_sum = sum(list_col))
```

To apply mutate functions to a list-column, you need to wrap the function you want to apply in a map function.

```{r}
tibble(list_col = list(c(1, 5, 7), 
                       5, 
                       c(10, 10, 11))) %>%
  mutate(list_sum = map(list_col, sum))
```

Since `map()` returns a list itself, the `list_sum` column is thus itself a list

```{r}
tibble(list_col = list(c(1, 5, 7), 
                       5, 
                       c(10, 10, 11))) %>%
  mutate(list_sum = map(list_col, sum)) %>% 
  pull(list_sum)
```

What could we do if we wanted it to be a vector? We could use the `map_dbl()` function instead!

```{r}
tibble(list_col = list(c(1, 5, 7), 
                       5, 
                       c(10, 10, 11))) %>%
  mutate(list_sum = map_dbl(list_col, sum))
```

### Nesting the gapminder data

Think of an individual data frame inside the nested data as `.x`. Again, I will first figure out the code for calculating the mean life expectancy for the first entry of the column. The following code defines `.x` to be the first entry of the data column (this is the data frame for Asia).

```{r}
# the first entry of the "data" column
.x <- gapminder_nested %>% pluck("data", 1)
.x
```

Then to calculate the average life expectancy for Asia, I could write

```{r}
mean(.x$lifeExp)

```

So copy-pasting this into the tilde-dot anonymous function argument of the `map_dbl()` function within `mutate()`, I get what I wanted!

```{r}
gapminder_nested %>% 
  mutate(avg_lifeExp = map_dbl(data, ~{mean(.x$lifeExp)}))
```

This code iterates through the data frames stored in the `data` column, returns the average life expectancy for each data frame, and concatenates the results into a numeric vector (which is then stored as a column called `avg_lifeExp`).

Hopefully it helped you understand why you need to wrap mutate functions inside map functions when applying them to list columns.

The next exampe will demonstrate how to fit a model separately for each continent, and evaluate it, all within a single tibble. First, I will fit a linear model for each continent and store it as a list-column. If the data frame for a single continent is `.x`, then the model I want to fit is `lm(lifeExp ~ pop + gdpPercap + year, data = .x)` (check for yourself that this does what you expect). So I can copy-past this command into the `map()` function within the `mutate()`


```{r}
# fit a model separately for each continent
gapminder_nested <- gapminder_nested %>% 
  mutate(lm_obj = map(data, ~lm(lifeExp ~ pop + gdpPercap + year, data = .x)))
gapminder_nested
```

Where the first linear model (for Asia) is

```{r}
gapminder_nested %>% pluck("lm_obj", 1)
```

I can then predict the response for the data stored in the data column using the corresponding linear model. So I have two objects I want to iterate over: the data and the linear model object. This means I want to use map2(). When things get a little more complicated I like to have multiple function arguments, so I’m going to use a full anonymous function rather than the tilde-dot shorthand.

```{r}
# predict the response for each continent
gapminder_nested <- gapminder_nested %>% 
  mutate(pred = map2(lm_obj, data, function(.lm, .data) predict(.lm, .data)))
gapminder_nested

```

And I can then calculate the correlation between the predicted response and the true response, this time using the `map2()_dbl` function since I want the output the be a numeric vector rather than a list of single elements.

```{r}
# calculate the correlation between observed and predicted response for each continent
gapminder_nested <- gapminder_nested %>% 
  mutate(cor = map2_dbl(pred, data, function(.pred, .data) cor(.pred, .data$lifeExp)))
gapminder_nested
```

## Additional purrr functionalities for lists

To demonstrate how to use purrr to manipulate lists, we will split the gapminder dataset into a *list of data frames* (which is kind of like the converse of a data frame containing a list-column). To make sure it’s easy to follow, we will only keep 5 rows from each continent.

```{r}
set.seed(23489)
gapminder_list <- gapminder %>% split(gapminder$continent) %>%
  map(~sample_n(., 5))
gapminder_list
```

### Keep/Discard: select_if for lists


`keep()` only keeps elements of a list that satisfy a given condition, much like `select_if()` selects columns of a data frame that satisfy a given condition.

The following code only keeps the gapminder continent data frames (the elements of the list) that have an average (among the sample of 5 rows) life expectancy of at least 70.

```{r}
gapminder_list %>%
  keep(~{mean(.x$lifeExp) > 70})
```

`discard()` does the opposite of `keep()`: it discards any elements that satisfy your logical condition.

### Reduce

`reduce()` is designed to combine (reduces) all of the elements of a list into a single object by iteratively applying a binary function (a function that takes two inputs)

For instance, applying a reduce function to add up all of the elements of the vector `c(1, 2, 3)` is like doing `sum(sum(1, 2), 3)`: first it applies `sum` to `1` and `2`, then it applies `sum` again to the output of `sum(1, 2)` and `3`

```{r}
reduce(c(1, 2, 3), sum)
```

`accumulate()` also returns the intermediate values.

```{r}
accumulate(c(1, 2, 3), sum)
```

An example of when `reduce()` might come in handy is when you want to perform many `left_join()`s in a row, or to do repeated `rbinds()` (e.g. to bind the rows of the list back together into a single data frame)

```{r}
gapminder_list %>%
  reduce(rbind)
```

### Logical statements for lists

Asking logical questions of a list can be done using `every()` and `some()`. For instance to ask whether every continent has average life expectancy greater than 70, you can use `every()`

```{r}
gapminder_list %>% every(~{mean(.x$life) > 70})
```

To ask whether some continents have average life expectancy greater than 70, you can use `some()`

```{r}
gapminder_list %>% some(~{mean(.x$life) > 70})

```

An equivalent of `%in%` for lists is `has_element()`.

```{r}
list(1, c(2, 5, 1), "a") %>% has_element("a")
```

Most of these functions also work on vectors.

